---
title: "CSC3007-Team-Black-2"
subtitle: "Add subtitle here"
author: "CHIA YI XUAN, TSUI SAU CHI, JEANIE OH JUN NING, VARADHARAJAN JAYAPRIYA, GU JINMING, WONG JUN HAO"
format:
  revealjs: 
    slide-number: true
---

```{r}
#| label: required-packages
#| message: false

library(tidyverse)
library(countrycode)
library(directlabels)
library(ggrepel)
library(lintr)
library(readxl)
library(scales)
```

```{r}

# Load & Clean the dataset / prepare it here.
```

# Relationship between nation's legal framework achievements in gender equality and woman's success in the economy.

## Purpose of This Presentation

## Research Questions

To what extent does a nation's legal framework on gender equality affect economic growth for women?

Specifically, what is the relationship between a [nation's legal framework achievements in gender equality]{.underline} and [woman's success in the economy]{.underline}?

Can [GDP]{.underline}, [Happy Planet Index Score]{.underline} and [Region]{.underline} serve as predictors for a nation's legal framework achievements in gender equality & proportion of women in managerial positions?

## Importance of This Research

-   Economic empowerment:

The economic empowerment of women has broader implications for overall economic development.

-   Policy formulation:

Inform policymakers and lawmakers about the effectiveness and relevance of legal frameworks in fostering economic growth for women.

## Importance of This Research

-   Gender equality:

The impact of these legal frameworks is crucial in advancing gender equality efforts.

-   Societal impact:

Sheds light on the potential benefits and consequences of legal frameworks on economic growth, impacting woman's lives and opportunities.

## UN Sustainable Development Goals

Women participation in the economy is related to Goals 5 and 8.

## Selected Indicators

**Indicator 5.1.1**: Legal frameworks that promote enforce and monitor gender equality (percentage of achievement 0 - 100) \-- Area 3: employment and economic benefits

**Indicator 5.5.2:** Proportion of women in managerial positions

**Indicator 8.5.1:** Average hourly earnings of employees by sex, age, occupation and persons with disabilities

**Indicator 8.5.2:** Unemployment rate by sex, age and persons with disabilities

## Choropleth Map:  **5.1.1**

```{r}

```

## Choropleth Map:  **5.5.2**

```{r}

```

## **5.1.1 vs 5.5.2 Anticipation**

We anticipate that the higher the nation's legal framework achievements in gender equality would result in a higher percentage of women holding managerial positions.

-   Because the chance of gender-based discrimination should be reduced as a result of the legal frameworks in place, this should result in better opportunities for women to hold managerial positions.

## **5.1.1 vs 5.5.2 Relation**

```{r}

```

## **5.1.1 vs 5.5.2 Relation data re-expression if needed**

```{r}

```

## Choropleth Map:  **8.5.1**

```{r}

```

## Choropleth Map:  **8.5.2**

```{r}

```

## **8.5.1 vs 8.5.2 Anticipation**

We anticipate that countries with lower unemployment rates for women should have higher average hourly earnings for women.

-   Because countries with lower unemployment rates have a better economic environment which should result in higher average income for their citizens

## **8.5.1 vs 8.5.2 Relation**

```{r}

```

## **8.5.1 vs 8.5.2 Relation data re-expression if needed**

```{r}

```

## **Table of Women Economical Success**

```{r}

```

## **5.1.1 vs 8.5.2 Anticipation**

We anticipate that having a higher the nation's legal framework achievements in gender equality would result in lower unemployment rate for women.

-   Because by decreasing gender discrimination for women, more companies would be willing to employ women.

## **5.1.1 vs 8.5.2 Relation**

```{r}

```

## **5.5.2 vs 8.5.1 Anticipation**

We anticipate that having more women in managerial positions will increase the average hourly earning for female employees.

-   Because in general, employees in a managerial position would receive higher income and with more women holding managerial positions, the average hourly earning for female employees will increase.

## **5.5.2 vs 8.5.1 Relation**

```{r}

```

## K-Means Clustering

## Pairs Plot Highlighting Clusters

```{r}

```

## Decision Tree

```{r}

# 3 country-level variable are used here.
```

## Evaluation of Decision Tree

## Summary of Quantitative Results

## Policy Implications
